High-throughput Genomic Paper Notes

### 2017-10-01 Paper_notes/Paper_notes.Rmd
Changed Lipshutz et al. summary to be below word count

Lipshutz et al. have developped a tool to collect analyze vast amounts of genetic and cellular information simultaneously from nucleic acid strands. Through 'fabrication of hundred of thousands of polynucleotides  at high spatial resolution on a precise location' and 'laser confocal fluorescence scanning', they were able to design a DNA probe array to get complmentary sequences. For example, the array is coated with a chemistry protecting group to prevent DNA deposition and then a mask is placed on top to expose specified regions. A light is used to knocked off exposed protecting group and then add certain solution of nucleitide incubations to hybridize to the nucleitide at that location. This cycle is repeated until a synthesized polynucleotides of about 300,000 are created at specific locations of the 1.28 cm $\times$ 1.28 cm array. The array that can contain approximately 40,000 human genes are used for expression monitoring to understand a gene function. From these arrays, fluorescence intensity image obtained from the array are compared with the reference sequence to be deemed perfect match (PM) and mismatch (MM). The difference of PM and MM can help reduce background noise and cross-hybridization.

This technique is useful as there is a need for 'monitoring expression levels of a large number of genes repeatedly, routinely, and reproducibily' that do not need physcial intermediaries such as cDNA, PCR products or clones and the process that come with it to prepare, verify and cataloque the large number of sequence. Again, for a single position on the DNA, one thousand sets of four probe pairs (1 complementary probe at the specific region and 3 similar complementary probes at anywhere but the specific position) are conducted to detect variants in DNA sequence - by identifing the difference (given a specific place) or positions. Using photolithography and light-activated chemistry, they are able to take advantage of the complementary properties of genes and the target design of the probes to monitor a large amount of expression levels. Companies such as Affymetrix are developping software tools to manage, monitor, genotype, and sequence analyze these huge datasets.

>>>>

Lipshutz et al. have developped a tool to collect analyze vast amounts of genetic and cellular information simultaneously from nucleic acid strands. Using photolithography and light-activated chemistry, they are able to take advantage of the complementary properties of genes and the target design of the probes, they were able to design a DNA probe array to obtain complementary sequences and monitor a large amount of expression levels. For example, the array is coated with a chemistry protecting group to prevent DNA deposition and then a mask is placed on top to expose specified regions. A light is used to knocked off exposed protecting group and then add certain solution of nucleitide incubations that hybridize to the nucleitide at that location. This cycle is repeated until a synthesized polynucleotides of about 300,000 are created at specific locations on the 1.28 cm $\times$ 1.28 cm array. The array that can contain approximately 40,000 human genes are used for expression monitoring to understand a gene function. Fluorescence intensity image of the array show perfect match (PM) and mismatch (MM) probe pairs. The difference of PM and MM can help reduce background noise and cross-hybridization in order to detect variants in DNA sequence - to identify the difference and the position of nucleitides. For a single position on the DNA, there can a number of PM/MM probe pairs.

This technique is useful as there is a need for 'monitoring expression levels of a large number of genes repeatedly, routinely, and reproducibily' that do not need physcial intermediaries such as cDNA, PCR products or clones and the process to prepare, verify and cataloque them. Companies such as Affymetrix are developping software tools to manage, genotype, and sequence analyze these datasets.
